MSB1015_Assignment 2

### 2019-09-28 MSB1015_Assignment2_SuzannetenHage.Rmd
Update MSB1015_Assignment2_SuzannetenHage.Rmd
I ran the same script, but this time normalized the data before training the models to see whether it improved the results. It did not.

data = query_wikidata('SELECT ?comp ?compLabel ?bp ?bpUnit ?bpUnitLabel ?smiles WHERE {   
    ?comp wdt:P31/wdt:P279* wd:Q41581 ;
    p:P2102 [ps:P2102 ?bp ;           
            psv:P2102/wikibase:quantityUnit  
            ?bpUnit         
            ] ;
    wdt:P233 ?smiles.
    SERVICE wikibase:label { bd:serviceParam wikibase:language "[AUTO_LANGUAGE],en". } 
    } '
)

@@

#Reformatting of smiles into correct format for CDK. Kekulise checks for electrons, and prevents parsing of incorrect smiles. 
parsed_smiles <- parse.smiles(data$smiles, kekulise=TRUE)

@@

#Find all rownumbers of the values in Celcius. 
celcius <-  which(grepl('degree Celsius', data$bpUnitLabel))
for (i in 1:length(celcius)) {
  #Convert the value to kelvin
  data$bp[celcius[i]] <- data$bp[celcius[i]] + 273.15
  #Convert the label to kelvin. 
  data$bpUnitLabel[celcius[i]] <- 'kelvin'
}

#Find all rownumbers of the values in Fahrenheit. 
fahrenheit <- which(grepl('degree Fahrenheit', data$bpUnitLabel))
for (i in 1:length(fahrenheit)) {
  #Convert the value to kelvin
  data$bp[fahrenheit[i]] <- (data$bp[fahrenheit[i]] + 459.67) * (5/9)
  #Convert the label to kelvin. 
  data$bpUnitLabel[fahrenheit[i]] <- 'kelvin'
}

@@

3b) To be able to test how well the model I train performs, it is necessary to keep a part of our data aside as testing data. In the following code you can easily change the size of the test data by adjusting percentage_test.

@@

#Splitting data into a training and test set. 
set.seed(50) 
sample = sample.split(full_data, SplitRatio = percentage_test)
train = subset(full_data, sample == FALSE)
test  = subset(full_data, sample == TRUE)

@@

#Create null_model
null_model <- mean(train$data.bp)

#Get error of null model
error_null_model <- rmse(test$data.bp, null_model)

@@

#Training of the PLS model. 
PLS_model <- plsr(data.bp ~ (apol + WPATH + WPOL), 
                  data = train, 
                  validation = "CV"
                  )

@@

#Plot the predicted data vs. the original data. 
plot(test$data.bp, predicted.bp[,,3], xlim=c(250,1000), ylim=c(250,1000))

>>>>

data_raw = query_wikidata('SELECT ?comp ?compLabel ?bp ?bpUnit ?bpUnitLabel ?smiles WHERE {   
    ?comp wdt:P31/wdt:P279* wd:Q41581 ;
    p:P2102 [ps:P2102 ?bp ;           
            psv:P2102/wikibase:quantityUnit  
            ?bpUnit         
            ] ;
    wdt:P233 ?smiles.
    SERVICE wikibase:label { bd:serviceParam wikibase:language "[AUTO_LANGUAGE],en". } 
    } '
)

@@

#Reformatting of smiles into correct format for CDK. Kekulise checks for electrons, and prevents parsing of incorrect smiles. 
parsed_smiles <- parse.smiles(data_raw$smiles, kekulise=TRUE)

@@

#Find all rownumbers of the values in Celcius. 
celcius <-  which(grepl('degree Celsius', data_raw$bpUnitLabel))
for (i in 1:length(celcius)) {
  #Convert the value to kelvin
  data_raw$bp[celcius[i]] <- data_raw$bp[celcius[i]] + 273.15
  #Convert the label to kelvin. 
  data_raw$bpUnitLabel[celcius[i]] <- 'kelvin'
}

#Find all rownumbers of the values in Fahrenheit. 
fahrenheit <- which(grepl('degree Fahrenheit', data_raw$bpUnitLabel))
for (i in 1:length(fahrenheit)) {
  #Convert the value to kelvin
  data_raw$bp[fahrenheit[i]] <- (data_raw$bp[fahrenheit[i]] + 459.67) * (5/9)
  #Convert the label to kelvin. 
  data_raw$bpUnitLabel[fahrenheit[i]] <- 'kelvin'
}

@@

3b) Normalization of the data.

```{r} 
#Normalization of data
data <- data.frame(descriptors, data_raw$bp)

#Devide every column by its maximum value. 
for (i in 1:ncol(data)) {
  data[,i] <- data[,i] * (1/max(data[,i]))
}

3c) To be able to test how well the model I train performs, it is necessary to keep a part of our data aside as testing data. In the following code you can easily change the size of the test data by adjusting percentage_test.

@@

#Splitting data into a training and test set. 
set.seed(50) 
sample = sample.split(data, SplitRatio = percentage_test)
train = subset(data, sample == FALSE)
test  = subset(data, sample == TRUE)

@@

#Create null_model
null_model <- mean(train$data_raw.bp)

#Get error of null model
error_null_model <- rmse(test$data_raw.bp, null_model)

@@

#Training of the PLS model. 
PLS_model <- plsr(data_raw.bp ~ (apol + WPATH + WPOL), 
                  data = train, 
                  validation = "CV"
                  )

@@

#Plot the predicted data vs. the original data. 
plot(test$data_raw.bp, predicted.bp[,,3])
